# Part A. Scraping data in table format with rvest and analysing it using the tidyverse (70 points)

__Name: Brooke Scoles

In this part of the assignment we will scrape tabular data, clean it after scraping, and analyse it. The following [Wikipedia site](https://en.wikipedia.org/wiki/COVID-19_pandemic_cases) contains a large table of _reported_ cumulative Covid 19 cases per month since January 2020 ("Monthly Totals/Cases by country"). 

Load all packages in the code cell below that you use in this part of the assignment:

```{r}
suppressMessages(library(rvest))
suppressMessages(library(tidyverse))
suppressMessages(library(lubridate))
```

1. Parse the HTML content of this URL "https://en.wikipedia.org/wiki/COVID-19_pandemic_cases" into R with the `rvest` package and its function `read_html()`. Extract all tables with `html_table()`. Print the length of the list of tables. (10 points)

```{r}
url_a <- "https://en.wikipedia.org/wiki/COVID-19_pandemic_cases"
html_a <- read_html(url_a) 
tables_a <- html_table(html_a, fill = TRUE)
length(tables_a)
```
**This list of tables contains 20 tables.**

2. This returned a large number of potential tables. Let us instead extract the main table in a more targeted way. For this determine the XPath of the "Monthly Totals/Cases by country" table (it is the table with around 220 rows) with your browser's inspect function. With this XPath, extract the respective node with `html_nodes()`. Print the length of the resulting list of nodes, it should now only contain one element. Afterwards, extract the table with `htm_table()` and the option fill=TRUE. (10 points)

```{r}
cases_by_country_table <- url_a %>%
  read_html() %>%
  html_nodes(xpath='//*[@id="mw-content-text"]/div[1]/table[3]') %>%
  html_table(fill=TRUE)
cases_by_country_table <- cases_by_country_table[[1]] #extracts the first element from the list
head(cases_by_country_table)
```
**Used strategy from lecture to extract table, then removed the blank row and printed the first few lines.**

3. Next, try to clean the scraped data and transform it into a `tidy` format. This is an important part of web scraping as data seldom can be obtained in its final form directly from websites.

3.1 The initial rows before the first country contain information that is not needed here. Delete these rows (i.e. an empty row, world-wide figures, days to double, and the countries and territories row which shows the amount of countries with positive cases). Leave all rows with countries in the table. Then rename the first column to "country" and drop the "First case" column as it is not needed in the following. (5 points)

3.2 Delete all commas from the columns with cases (i.e. all columns that represent months)  and convert them into numeric format. (5 points)

3.3 These columns with cases actually contain a date variable just like the example we have seen in the last weeks. Transform the data into a `tidy` long format with the function `pivot_longer()`. This should yield a new "date" and a new "cases" column. Print out only the head of your dataframe with the function `head()`. This final long dataframe should only have three columns: country, date, and cases.  (10 points)

```{r}
#3.1
cases_by_country_table <- cases_by_country_table[-c(1:4),-c(2)] #Drops the first rows and the first case column
colnames(cases_by_country_table)[colnames(cases_by_country_table) == "Date"] <- "Country"
#3.2 delete commas and convert to numeric
for(i in 2:11){
  cases_by_country_table[,i] <-  as.numeric(gsub(",", "", cases_by_country_table[,i] )) #followed method from the lecture
 }

#3.3
#cases_by_country_table_test <- cases_by_country_table
cases_by_country_table <- cases_by_country_table %>%
  pivot_longer(
    cols = -Country,
    names_to = c('Date'),
    #names_prefix = "X",
    #names_sep = '.Q',
    values_to = "Cases"
  ) 
head(cases_by_country_table)

```
**First I removed the rows that were not needed. Then, I removed commas and converted to numeric for the columns for the months. To do this, I created a for loop for those columns (columns 2-11) and iterated over each column, converting each to numeric. Lastly, used pivot_longer() to convert the table into long format.**

4. Transform the content of the date column into a datetime object with the `lubridate` package. Similar cases are discussed here: https://r4ds.had.co.nz/dates-and-times.html (10 points)

```{r}
cases_by_country_table$Year <- '2020'

#separate the column by whitespace
cases_by_country_table<- separate(data = cases_by_country_table, col = Date, into = c("Month", "Day"))
cases_by_country_table$Month <- match(cases_by_country_table$Month,month.abb)
cases_by_country_table$FullDate <-
  paste(cases_by_country_table$Year,cases_by_country_table$Month,cases_by_country_table$Day,sep="/")
cases_by_country_table$FullDate2 <- as.Date(cases_by_country_table$FullDate, format = "%Y/%m/%d")

#remove additional columns
cases_by_country_table <- cases_by_country_table[,-c(2,3,5,6)]
colnames(cases_by_country_table)[colnames(cases_by_country_table) == "FullDate2"] <- "Date"
cases_by_country_table <- cases_by_country_table[c(1,3,2)]
head(cases_by_country_table)

```


5. Next, the goal is to add a fourth variable which contains the percentage change of cases:

5.1 Sort the long dataframe by country name and by date with the `arrange()` function. The first observation should be Afghanistan in January. (5 points)

5.2 Use the `mutate()` function to create a new column with the name "cases_percentage_change" which contains the monthly percentage changes in cases. During or after this transformation make sure to set the first percentage change for each country to NA. There cannot be a percentage change for this month as we do not have a value for December 2019. (5 points)

5.3 In addition to the NA values, the resulting percentage change columns contains NaN and Inf values. Explain (just in words, no code needed) how NaN and Inf each resulted here. (5 points) 


```{r}
#5.1
cases_by_country_table <- arrange(cases_by_country_table, Country, Date)
head(cases_by_country_table)
#5.2
cases_by_country_table <- cases_by_country_table %>%
  group_by(Country) %>%
  mutate(cases_percentage_change = round((Cases - lag(Cases)) / lag(Cases) * 100, digits = 2) )
head(cases_by_country_table)
```
**In 5.1 above you can see the result of rearranging the table. Then the percentage change was calculated in the fourth column. NAN result from dividing by zero by zero. There is no percentage change from zero to zero. Inf result from dividing by zero because the percentage change from zero to one is infinity.**

6. Out of all countries that had at least 100,000 reported cumulative cases in October, which 10 countries had the highest percentage increase __from the beginning of September to the beginning of October__? By what percentage did their reported cases rise in this last month? (5 points)

```{r}
top_ten <- cases_by_country_table %>%
  filter(Date == "2020-10-01",
         Cases >= 100000) %>%
  arrange(-cases_percentage_change)
head(top_ten, 10)
```

